Amphistomy increases leaf photosynthesis more in coastal than montane plants of Hawaiian ʻilima (*Sida fallax*)

### 2023-10-29 ms/ms.qmd
added ags-curve example figure

We cut thin sections using two razor blades taped together. We sectioned the leaf in a petri dish of water, wet-mounted sections onto a slide, and took digital micrographs using a brightfield microscope, as described above.

@@

At each site, we selected one representative leaf from one plant near the middle of the transect for gas exchange measurements using a portable infrared gas analyzer (LI-6800PF, LI-COR Biosciences, Lincoln, NE, USA). We estimated the photosynthetic rate ($A$) and stomatal conductance to water vapor ($g_\text{sw}$) at saturating light ($\text{photosynthetic photon flux density (PPFD)} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), ambient CO$_2$ (415 ppm), and $T_\mathrm{leaf} = 25.0 \textendash 29.3° \text{C}$. We also estimated 'amphistomy advantage' ($\mathrm{AA}$) *sensu* @parkhurst_adaptive_1978, but with modification. For each leaf, we measured the photosynthetic rate of an untreated amphistomatous leaf ($A_{\mathrm{amphi}}$) over a range of $g_\text{sw}$ values. We refer to this as an $A \textendash g_\text{sw}$ curve. We compared the $A \textendash g_\text{sw}$ curve of the untreated leaf to the photosynthetic rate of pseudohypostomatous leaf ($A_\mathrm{hypo}$), which is the same leaf but with gas exchange through the upper surface blocked by a neutral density plastic (propafilm). Hypostomy refers to leaves with stomata only present on the lower, typically abaxial, surface. We refer to the untreated and partially blocked leaves as "amphi" and "pseudohypo", respectively. $\mathrm{AA}$ is calculated as the log-response ratio of $A$ compared at the same total $g_\text{sw}$:

@@

To control for order effects, we alternated between starting with amphi (untreated) or pseudohypo (treated) leaf measurements. In the field, we acclimated the focal leaf to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), high relative humidity ($\mathrm{RH} = 70\%$), $T_\text{leaf} = 25-30~° \text{C}$, depending on ambient temperature, until $A$ and $g_\text{sw}$ reach their maximum. We used these data as our estimates of maximum $A$ and $g_\text{sw}$. After that, we decreased $\mathrm{RH}$ to $\approx 10\%$ to induce rapid stomatal closure without biochemical downregulation. Hence, $A_\text{amphi}$ and $A_\text{hypo}$ were both measured at low chamber humidity after the leaf had acclimated to high humidity. All other environmental conditions in the leaf chamber remained the same. We logged data until $g_\text{sw}$ reached its nadir. We then repeated the process with the other treatment (amphi or pseudohypo). We did not detect an effect of treatment order on $\mathrm{AA}$ (results not shown).

@@

It is not feasible to record $A_\mathrm{amphi}$ and $A_\mathrm{hypo}$ at the exact same $g_\text{sw}$. To overcome this, we fit $A \textendash g_\text{sw}$ curves using a linear regression of $\text{log}(g_\mathrm{sw})$ on $A$ to interpolate modeled $A$ for amphi and pseudohypo leaves at the same $g_\text{sw}$. In 10 of 12 leaves, the minimum $g_\text{sw}$ of the amphi curve was smaller than the maximum $g_\text{sw}$ of the pseudohypo curve (i.e. the curves overlapped for a range of $g_\text{sw}$ values). In those cases, we estimated $\mathrm{AA}$ at the $g_\mathrm{sw}$ value in the middle of the range of overlap. In 2 of 12 leaves, the $A \textendash g_\text{sw}$ curves did not quite overlap because the minimum $g_\text{sw}$ of the amphi curve was slightly greater than the maximum $g_\text{sw}$ of the pseudohypo curve. In those cases, we estimated $\mathrm{AA}$ by extrapolating slightly, `r gsw_extrap`, beyond the measured curves to the $g_\mathrm{sw}$ value in between the curves. We fit Bayesian regressions using the *R* package **brms** version `r fit_licor$version$brms` [@burkner_brms_2017] with MCMC sampling in *Stan* [@stan_development_team_stan_2023]. We used CmdStan version `r fit_licor$version$cmdstan` and **cmdstanr** version `r fit_licor$version$cmdstanr` [@gabry_cmdstanr_2023] to interface with *R* version `r getRversion()` [@r_core_team_r:_2023]. We sampled the posterior distribution from `r dim_fit_licor[2]` chains with `r dim_fit_licor[1]` iterations each after `r dim_fit_licor[1]` warmup iterations. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. The key prediction is that $\mathrm{AA}_\text{coastal} > \mathrm{AA}_\text{montane}$, meaning the 95\% confidence intervals of $\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane}$ should be positive and not encompass 0.

@@

We tested whether leaf thickness (log-transformed) varied between coastal and montane populations and among individuals within populations using a Bayesian mixed-effects model with habitat as a fixed effect and individual plant and site as random effects. We used the *R* package **brms** version `r fit_thickness$version$brms` [@burkner_brms_2017] to fit the model in *Stan* [@stan_development_team_stan_2023] with CmdStan version `r fit_thickness$version$cmdstan` and **cmdstanr** version `r fit_thickness$version$cmdstanr` [@gabry_cmdstanr_2023]. We sampled the posterior distribution from `r dim_fit_thickness[2]` chains with `r dim_fit_thickness[1]` iterations each after `r dim_fit_thickness[1]` warmup iterations. We estimated the relationship between population average leaf thickness and $\mathrm{AA}$ measured from a single individual per population. We used this approach most of the variation in leaf thickness was among sites and the plant selected for gas exchange measurements was not always among the plants randomly selected for leaf thickness, precluding individual level correlation. We propagated uncertainty about in $\mathrm{AA}$ and leaf thickness by integrating over the entire posterior distribution sample for each variable. The key prediction is that the effect of leaf thickness on $\mathrm{AA}$ is positive, meaning the 95\% confidence interval of the slope should be positive and not encompass 0.

@@

We tested whether $g_\mathrm{smax,ratio}$ varied between coastal and montane populations and among individuals within populations using a Bayesian mutliresponse, mixed-effects model. The modeled response variables are stomatal count and guard cell length on each surface. Counts were modeled as negative binomially distributed variable from a latent stomatal density and a parameter $\phi$ to estimate overdispersion in counts relative to a Poisson model. For all traits, the explanatory variables were habitat as a fixed effect and leaf within individual plant, individual plant, and site as random effects. We used the *R* package **brms** version `r fit_stomata$version$brms` [@burkner_brms_2017] to fit the model in *Stan* [@stan_development_team_stan_2023] with CmdStan version `r fit_stomata$version$cmdstan` and **cmdstanr** version `r fit_stomata$version$cmdstanr` [@gabry_cmdstanr_2023]. We interpolated missing adaxial guard cell lengths from `r length(which(fit_stomata$data$upper_number_of_stomata == 0))` out of `r nrow(fit_stomata$data)` samples with zero adaxial stomata using the "mi" function in **brms** package. We sampled the posterior distribution from `r dim_fit_stomata[2]` chains with `r dim_fit_stomata[1]` iterations each after `r dim_fit_stomata[1]` warmup iterations. From each posterior sample, we calculated $g_{\mathrm{smax,ratio}}$ as:

$$g_\mathrm{smax,ratio} = \frac{g_\mathrm{smax,upper}}{g_\mathrm{smax,lower} + g_\mathrm{smax,upper}},$$ where $g_\mathrm{smax,lower}$ and $g_\mathrm{smax,upper}$ are maximum stomatal conductance to water vapor at $T_\mathrm{leaf} = 25 ^ \circ$ C on the lower and upper surface, respectively. The maximum stomatal conductance is calculated from stomatal density and length, assuming that stomata are fully open, following @sack_developmental_2016.

<!-- They defined a biophysical and morphological constant as: -->

@@

We estimated the relationship between leaf $g_\mathrm{smax,ratio}$ and $\mathrm{AA}$ measured from a single leaf per population. We propagated uncertainty about in $\mathrm{AA}$ and $g_\mathrm{smax,ratio}$ by integrating over the entire posterior distribution sample for each variable. The key prediction is that the effect of $g_\mathrm{smax,ratio}$ on $\mathrm{AA}$ is positive until $g_\mathrm{smax,ratio} < 0.5$.

@@

Amphistomy increases photosynthesis in leaves of coastal ʻilima plants more than that of montane plants. $\mathrm{AA}$ was significantly greater than 0 (95% confidence intervals did not overlap 0) in 5 of 6 coastal leaves, but only 1 of 6 montane leaves (\autoref{fig:habitat-aa}; see \autoref{fig:licor} for individual curves). Overall, the average $\mathrm{AA}$ among coastal and montane leaves is $`r AA_coastal`$ and $`r AA_montane`$, respectively; the difference in average $\mathrm{AA}$ between habitat types is $\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane} = `r AA_diff`$. Posterior predictions closely match observed values of $A$ (\autoref{fig:pp-licor}), indicating an adequate model fit from which we can interpolate between measurements reliably. It also suggests that slight extrapolation beyond the data should be reliable, but this is less certain. When we remove two leaves where extrapolated slightly beyond fitted $A \textendash g_\mathrm{sw}$ curves, we estimate that $\mathrm{AA}_\text{coastal}$ is still positive, $`r AA_coastal1`$, but the difference between coastal and montane leaves is smaller,  $`r AA_diff1`$, and confidence intervals slightly overlap 0.  Maximum photosynthetic rate was slightly, but not significantly higher in coastal leaves (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['A']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['A']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['A']][['p.value']], 2)`$); total stomatal conductance was similar (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['gsw']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['gsw']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['gsw']][['p.value']], 2)`$) in coastal and montane leaves (\autoref{fig:habitat-Ags}).

@@

<!-- add tom buckley as coauthor or acknowledge help developing method -->

The authors thank Kasey Barton for feedback on an earlier version of this manuscript, the University of Hawaiʻi honors council for guidance to GT, Tawn Speetjens for access to one montane site, startup funds from the University of Hawaiʻi, NSF Award 1929167 to CDM, and NSF Award 2307341 to TNB.

@@

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thetable{S\arabic{table}}    
\renewcommand\theequation{S\arabic{equation}}    
\setcounter{figure}{0}    
\setcounter{table}{0}    
\setcounter{equation}{0}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:licor}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:pp-licor}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:habitat-Ags}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:habitat-gmaxratio}
\end{figure}
```
```{=tex}
\begin{table}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{tab:sum-data}
\end{table}
```

>>>>

We cut thin sections using two razor blades taped together. We sectioned the leaf in a petri dish of water, wet-mounted sections onto a slide, and took digital micrographs using a brightfield microscope, as described above. Leaf thickness is measured as the length from upper cuticle to lower cuticle.

@@

At each site, we selected one representative leaf from one plant near the middle of the transect for gas exchange measurements using a portable infrared gas analyzer (LI-6800PF, LI-COR Biosciences, Lincoln, NE, USA). We estimated the photosynthetic rate ($A$) and stomatal conductance to water vapor ($g_\text{sw}$) at saturating light ($\text{photosynthetic photon flux density (PPFD)} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), ambient CO$_2$ (415 ppm), and $T_\mathrm{leaf} = 25.0 \textendash 29.3° \text{C}$. The midday irradiance in coastal ʻilima typically meets or even exceeds a PPFD of $2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$ and previous experiments with sun leaves revealed that $2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$ is always at or near saturating irradiance. Even though lower irradiance may be saturating for montane leaves, we used this higher value for all leaves to standardize conditions.

We also estimated 'amphistomy advantage' ($\mathrm{AA}$) *sensu* @parkhurst_adaptive_1978, but with modification. For each leaf, we measured the photosynthetic rate of an untreated amphistomatous leaf ($A_{\mathrm{amphi}}$) over a range of $g_\text{sw}$ values (see below for details). We refer to this as an $A \textendash g_\text{sw}$ curve. We compared the $A \textendash g_\text{sw}$ curve of the untreated leaf to the photosynthetic rate of pseudohypostomatous leaf ($A_\mathrm{hypo}$), which is the same leaf but with gas exchange through the upper surface blocked by a neutral density plastic (propafilm). Hypostomy refers to leaves with stomata only present on the lower, typically abaxial, surface. We refer to the untreated and partially blocked leaves as "amphi" and "pseudohypo", respectively. $\mathrm{AA}$ is calculated as the log-response ratio of $A$ compared at the same total $g_\text{sw}$:

@@

\autoref{fig:ags-curve} illustrates our method for collecting $A \textendash g_\text{sw}$ curves. We collected two curves per leaf, an amphi (untreated) curve and a pseudohypo (treated) curve. To control for order effects, we alternated between starting with amphi or pseudohypo leaf measurements. In the field, we acclimated the focal leaf to saturating light and high relative humidity ($\mathrm{RH} = 70\%$), as described above, until $A$ and $g_\text{sw}$ reach their maximum. We used these data as our estimates of maximum $A$ and $g_\text{sw}$. After that, we decreased $\mathrm{RH}$ to $\approx 10\%$ to induce rapid stomatal closure without biochemical downregulation. Hence, $A_\text{amphi}$ and $A_\text{hypo}$ were both measured at low chamber humidity after the leaf had acclimated to high humidity. All other environmental conditions in the leaf chamber remained the same. We logged data until $g_\text{sw}$ reached its nadir. We then repeated the process of acclimating the leaf to 70\% $\mathrm{RH}$ and inducing stomatal closure with low $\mathrm{RH}$ with the other treatment (amphi or pseudohypo). We did not detect an effect of treatment order on $\mathrm{AA}$ (results not shown).

@@

It is not feasible to record $A_\mathrm{amphi}$ and $A_\mathrm{hypo}$ at the exact same $g_\text{sw}$. To overcome this, we fit $A \textendash g_\text{sw}$ curves using a linear regression of $\text{log}(g_\mathrm{sw})$ on $A$ to interpolate modeled $A$ for amphi and pseudohypo leaves at the same $g_\text{sw}$. In 10 of 12 leaves, the minimum $g_\text{sw}$ of the amphi curve was smaller than the maximum $g_\text{sw}$ of the pseudohypo curve (i.e. the curves overlapped for a range of $g_\text{sw}$ values). In those cases, we estimated $\mathrm{AA}$ at the $g_\mathrm{sw}$ value in the middle of the range of overlap. In 2 of 12 leaves, the $A \textendash g_\text{sw}$ curves did not quite overlap because the minimum $g_\text{sw}$ of the amphi curve was slightly greater than the maximum $g_\text{sw}$ of the pseudohypo curve. In those cases, we estimated $\mathrm{AA}$ by extrapolating slightly, `r gsw_extrap`, beyond the measured curves to the $g_\mathrm{sw}$ value in between the curves. We fit Bayesian regressions using the *R* package **brms** version `r fit_licor$version$brms` [@burkner_brms_2017] with MCMC sampling in *Stan* [@stan_development_team_stan_2023]. We used CmdStan version `r fit_licor$version$cmdstan` and **cmdstanr** version `r fit_licor$version$cmdstanr` [@gabry_cmdstanr_2023] to interface with *R* version `r getRversion()` [@r_core_team_r:_2023]. We sampled the posterior distribution from `r dim_fit_licor[2]` chains with `r dim_fit_licor[1]` iterations each after `r dim_fit_licor[1]` warmup iterations per chain. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. The key prediction is that $\mathrm{AA}_\text{coastal} > \mathrm{AA}_\text{montane}$, meaning the 95\% confidence intervals of $\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane}$ should be positive and not encompass 0.

@@

We tested whether leaf thickness (log-transformed) varied between coastal and montane populations and among individuals within populations using a Bayesian mixed-effects model with habitat as a fixed effect and individual plant and site as random effects. We used the *R* package **brms** version `r fit_thickness$version$brms` [@burkner_brms_2017] to fit the model in *Stan* [@stan_development_team_stan_2023] with CmdStan version `r fit_thickness$version$cmdstan` and **cmdstanr** version `r fit_thickness$version$cmdstanr` [@gabry_cmdstanr_2023]. We sampled the posterior distribution from `r dim_fit_thickness[2]` chains with `r dim_fit_thickness[1]` iterations each after `r dim_fit_thickness[1]` warmup iterations per chain. We estimated the relationship between population average leaf thickness and $\mathrm{AA}$ measured from a single individual per population. We used this approach because most of the variation in leaf thickness was among sites and the plant selected for gas exchange measurements was not always among the plants randomly selected for leaf thickness, precluding individual level correlation. We propagated uncertainty about in $\mathrm{AA}$ and leaf thickness estimates by integrating over the entire posterior distribution sample for each variable. The key prediction is that the effect of leaf thickness on $\mathrm{AA}$ is positive, meaning the 95\% confidence interval of the slope should be positive and not encompass 0.

@@

We tested whether $g_\mathrm{smax,ratio}$ varied between coastal and montane populations and among individuals within populations using a Bayesian mutliresponse, mixed-effects model. The modeled response variables are stomatal count and guard cell length on each surface. Counts were modeled as negative binomially distributed variable from a latent stomatal density and a parameter $\phi$ to estimate overdispersion in counts relative to a Poisson model. For all traits, the explanatory variables were habitat as a fixed effect and leaf within individual plant, individual plant, and site as random effects. We used the *R* package **brms** version `r fit_stomata$version$brms` [@burkner_brms_2017] to fit the model in *Stan* [@stan_development_team_stan_2023] with CmdStan version `r fit_stomata$version$cmdstan` and **cmdstanr** version `r fit_stomata$version$cmdstanr` [@gabry_cmdstanr_2023]. We interpolated missing adaxial guard cell lengths from `r length(which(fit_stomata$data$upper_number_of_stomata == 0))` out of `r nrow(fit_stomata$data)` samples with zero adaxial stomata using the "mi" function in **brms** package. We sampled the posterior distribution from `r dim_fit_stomata[2]` chains with `r dim_fit_stomata[1]` iterations each after `r dim_fit_stomata[1]` warmup iterations per chain. From each posterior sample, we calculated $g_{\mathrm{smax,ratio}}$ as:

$$g_\mathrm{smax,ratio} = \frac{g_\mathrm{smax,upper}}{g_\mathrm{smax,lower} + g_\mathrm{smax,upper}},$$

where $g_\mathrm{smax,lower}$ and $g_\mathrm{smax,upper}$ are maximum stomatal conductance to water vapor at $T_\mathrm{leaf} = 25 ^ \circ$ C on the lower and upper surface, respectively. The maximum stomatal conductance was calculated from stomatal density and length, assuming that stomata are fully open, following @sack_developmental_2016:

$$g_\mathrm{smax} = bmds^{0.5}.$$
In this equation, $b$ is a biophysical constant, $m$ is a morphological constant, $d$ is the stomatal density, and $s$ is the stomatal complex area. We assume that $b$, which is determined by the molecular species, temperature, and air pressure, is the same for both surfaces; we assume that $m$, which is determined by guard cell allometry is also the same for both surfaces. Hence, the $b$ and $m$ constants cancel out of $g_\mathrm{smax,ratio}$ and only density and length ($l$), which is proportional to the square root of area, affect the ratio: $g_\mathrm{smax} \propto dl$.

@@

We estimated the relationship between leaf $g_\mathrm{smax,ratio}$ and $\mathrm{AA}$ measured from a single leaf per population. We propagated uncertainty about $\mathrm{AA}$ and $g_\mathrm{smax,ratio}$ by integrating over the entire posterior distribution sample for each variable. The key prediction is that the effect of $g_\mathrm{smax,ratio}$ on $\mathrm{AA}$ is positive until $g_\mathrm{smax,ratio} < 0.5$.

@@

Amphistomy increases photosynthesis in leaves of coastal ʻilima plants more than that of montane plants. $\mathrm{AA}$ was significantly greater than 0 (95% confidence intervals did not overlap 0) in 5 of 6 coastal leaves, but only 1 of 6 montane leaves (\autoref{fig:habitat-aa}; see \autoref{fig:licor} for individual curves). Overall, the average $\mathrm{AA}$ among coastal and montane leaves is $`r AA_coastal`$ and $`r AA_montane`$, respectively; the difference in average $\mathrm{AA}$ between habitat types is $\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane} = `r AA_diff`$. Posterior predictions closely match observed values of $A$ (\autoref{fig:pp-licor}), indicating an adequate model fit from which we can interpolate between measurements reliably. It also suggests that slight extrapolation beyond the data should be reliable, but this is less certain. When we remove two leaves where extrapolated slightly beyond fitted $A \textendash g_\mathrm{sw}$ curves, we estimate that $\mathrm{AA}_\text{coastal}$ is still positive, $`r AA_coastal1`$, but the difference between coastal and montane leaves is smaller,  $`r AA_diff1`$, and confidence intervals slightly overlap 0.  Maximum photosynthetic rate was slightly, but not significantly higher in coastal leaves (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['A']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['A']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['A']][['p.value']], 2)`$); total stomatal conductance was similar (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['gsw']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['gsw']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['gsw']][['p.value']], 2)`$) in coastal and montane leaves (\autoref{fig:habitat-Ags}). Water-use efficiency ($A / g_\mathrm{sw}$) was significantly higher in coastal leaves (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['wue']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['wue']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['wue']][['p.value']], 2)`$).

@@

The authors thank Kasey Barton for feedback on an earlier version of this manuscript, the University of Hawaiʻi honors council for guidance to GT, Tawn Speetjens for access to one montane site, TM Perez for advice on leaf sectioning, startup funds from the University of Hawaiʻi, NSF Award 1929167 to CDM, and NSF Award 2307341 to TNB.

@@

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thetable{S\arabic{table}}    
\renewcommand\theequation{S\arabic{equation}}    
\setcounter{figure}{0}    
\setcounter{table}{0}    
\setcounter{equation}{0}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:ags-curve}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:licor}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:pp-licor}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:habitat-Ags}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:habitat-gmaxratio}
\end{figure}
```
```{=tex}
\begin{table}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{tab:sum-data}
\end{table}
```

### 2023-10-29 ms/si.qmd
added ags-curve example figure

>>>>

```{=tex}
\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}
```

### 2023-10-29 ms/si.tex
added ags-curve example figure

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, interior hidden, enhanced, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

>>>>

\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}
